---
title: "Delaware River PIT tag data analysis"
author: "Ben Letcher"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

library(tidyverse)
library(lubridate)
library(knitr)
library(targets)
library(tarchetypes)
library(MCMCvis)
library(htmltools)

library(bs4Dash)
library(DT)
library(gt)
library(pingr)
library(shinybusy)
library(shinyWidgets)

options(tinytex.verbose = TRUE)
```

*Note*: the results shown here are preliminary and have not been officially reviewed by USGS, NYDEC or PA Fish and Boat

This notebook uses [targets](https://books.ropensci.org/targets/) to manage running code and updating R objects. Targets sets up dependencies among specified objects and only re-runs code as necessary (when an upstream component gets updated). This can save run times for projects with models that take a while to run, like capture-mark-recapture models.

Data preparation and model running happens using targets and exploration of the data and model runs is below in this Markdown document.

'tar_make()' runs all the R scripts and functions specified in '\_targets.R'. Only updated code or sections that are downstream from updated data are re-run.\
'tar_read()' reads 'target' data into the global environment.

To set up a targets project, use use_targets()

This section (tar_make()) reruns the model and has some other helpful functions.

```{r tar_make}
#  tar_watch(seconds = 10, outdated = FALSE, targets_only = TRUE)


# comment this out when knitting - get Latex error that it can't find the check mark the tar_make() uses
#tar_make()

# tar_prune() # cleans unused data files
#tar_invalidate(everything())
#tar_invalidate(ends_with("ttt"))


#str(d)
```

Load raw data for exploration\
Without a postscript ('\_main' or '\_trib'), the data are for all fish (main and trib).

```{r getTargets}
dRaw0 <- tar_read(dRaw0) #all data - including untagged
dRaw <- tar_read(dRaw) #all data for CMR models
d <- tar_read(target_d)
eh <- tar_read(target_eh)

```

Visualize the network - does not work with pdf output

```{r}
#tar_visnetwork()
```

Which rivers (Water) riverN corresponds to

```{r riverN}
table(d$Water, d$riverN)
```

### Raw data summary tables

```{r basic summaries}

kable(data.frame(ftable(d$date)))
#kable(data.frame(ftable(d$Water, d$riverN)))

#kable(data.frame(ftable(d$Water, d$riverN, d$date)))
kable(data.frame(ftable(d$species)))

### Number of unique tags
length(unique(d$tag))
```

Group observations by month.\
Luckily, sampling periods do not span months, so we can use month as a grouping variable for sampling occasion

```{r by month}
kable(data.frame(ftable(d$dateYM)))

```

### Tag information

Grouped by Water (sampling area)

```{r tag information}

tagN <- d %>% 
  group_by(tag, Water) %>%
  summarize(n = n()) %>%
  filter(tag != "") %>%
  arrange(desc(n))

### Number of times individual fish were observed
table(tagN$n)

### Number of times individual fish were observed by river
(table(tagN$Water, tagN$n))



```

Grouped by state

States\
*River* *size1* *size2* *size3*\
Main 1 2 3\
Trib 4 5 6

```{r tag information state}

tagN_s <- d %>% 
  group_by(tag, state) %>%
  summarize(n = n()) %>%
  filter(tag != "") %>%
  arrange(desc(n))

### Number of times individual fish were observed
table(tagN_s$n)

### Number of times individual fish were observed by river
table(tagN_s$state, tagN_s$n)
```

#### Grouped by main/trib

This is what is used for the primary analysis

```{r tag information mainTrib}

tagN_mt <- d %>% 
  group_by(tag, mainTrib) %>%
  summarize(n = n()) %>%
  filter(tag != "") %>%
  arrange(desc(n))

### Number of times individual fish were observed
table(tagN_mt$n)

### Number of times individual fish were observed by river
(table(tagN_mt$mainTrib, tagN_mt$n))


```

### Basic summary plots of raw tagging data

```{r basic graphs}

ggplot(d %>% filter(species == "brown trout"), aes(Length)) +
  geom_freqpoly() +
  geom_vline(xintercept = c(tar_read(target_sizeCutoff1), tar_read(target_sizeCutoff2))) +
  facet_grid(~mainTrib)

```

```{r data for TAME, ignore=TRUE}

# dTame <- d %>% 
#             select(Latitude, Longitude, tag, dateTime, species, Length, Weight) %>%
#             filter(tag != "", tag != "ad")
#             
# write.csv(dTame, './dataOut/dTame.csv', row.names = FALSE)

```

### Encounter histories

This is the data structure for the capture-recapture models. Each column is a sampling 'occasion' (here = month) and each row is an individual, where a '1' indicates capture and a '0' indicates not captured.

```{r eh}

str(eh$eh)
kable(head(eh$eh,8))

table(paste(eh$first, eh$last, sep="_"))
```

There are very few transitions between main and trib and we can characterize size transitions better with a growth model - run phiT_pT_psiT models separately for main and trib

```{r}
s <- eh$stateMatrix
s[s == 0] <- NA
  
s2 <- data.frame(s) %>%
  unite("all", sep = "", na.rm = TRUE, remove = TRUE) %>%
  #distinct() %>%
  arrange()

table(s2)
```

Summary info for years and occasions

```{r}

years <- colnames(eh$eh) %>%
  substr(6,9) %>%
  as.numeric()

months <- colnames(eh$eh) %>%
  substr(11,12) %>%
  as.numeric()

occs <- colnames(eh$eh)
```

## Models

'phi' = apparent survival (probability of staying in the area = p(survival) + p(not moving out of area)).\
'p' = probability of capture given that the fish is alive. 'psi' = probability of transitioning from one state to another. Here, states are size bins.

### phiT_pT_psiT_main

Load data 'main' for analysis

```{r getTargets main}
#d_tt <- tar_read(target_d_trib)
eh_main <- tar_read(target_eh_main)

str(eh_main$eh)
#kable(eh_main$eh[1:8,1:10])

table(paste(eh_main$first, eh_main$last, sep="_"))

```

```{r tt main, eval=TRUE}
### Read the model run into global memory
if (tar_exist_objects(c("tt_modelOut_main"))) {
  mod_tt_main <- tar_read(tt_modelOut_main)
  
  MCMCplot(object = mod_tt_main$mcmc, params = "betaPhi")
  MCMCplot(object = mod_tt_main$mcmc, params = "betaP")
  
   priors <- rnorm(tar_read(tt_runData_main)$nIter * tar_read(tt_runData_main)$nChains, 0, 1/sqrt(.1))
   MCMCtrace(object = mod_tt_main$mcmc,
             ISB = FALSE,
             exact = TRUE,
             params = c("betaPhi[1, 1]", "betaPhi[2, 1]", "betaPhi[3, 1]",
                        "betaPhi[1, 2]", "betaPhi[2, 2]", "betaPhi[3, 2]"
                        #"betaPhi[1, 3]", "betaPhi[2, 3]", "betaPhi[3, 3]"
                        ),
             pdf = FALSE,
             priors = priors
             )

      MCMCtrace(object = mod_tt_main$mcmc,
             ISB = FALSE,
             exact = TRUE,
             params = c("betaP[1, 1]", "betaP[2, 1]", "betaP[3, 1]",
                        "betaP[1, 2]", "betaP[2, 2]", "betaP[3, 2]"
                        #"betaP[1, 3]", "betaP[2, 3]", "betaP[3, 3]"
                        ),
             pdf = FALSE,
             priors = priors
             )
  
    modSummary_tt_main <- MCMCsummary(object = mod_tt_main$mcmc, round = 3) %>%
      rename(lo = '2.5%', med = '50%', hi = '97.5%')
}
```

### Main, phi, p and psi

```{r}
  nS <- mod_tt_main$myConstants$nStates
  nT <- mod_tt_main$myConstants$T 
    
  phi_tt_main <- modSummary_tt_main %>% 
    filter(substr(row.names(modSummary_tt_main), 1, 7) == "betaPhi") %>%
    add_column(data.frame(state = 1:nS, dateYM = rep(1:(nT - 1), each = nS))) %>%
    mutate(year = years[dateYM],
           month = months[dateYM], 
           river = "main")
  
  p_tt_main <- modSummary_tt_main %>% 
    filter(substr(row.names(modSummary_tt_main), 1, 6) == "betaP[") %>%
    add_column(data.frame(state = 1:nS, dateYM = rep(1:(nT - 1), each = nS))) %>%
    mutate(year = years[dateYM],
           month = months[dateYM], 
           river = "main")

  psi_tt_main <- modSummary_tt_main %>% 
    filter(substr(row.names(modSummary_tt_main), 1, 3) == "psi") %>%
    add_column(data.frame(stateFrom = 1:nS, stateTo = rep(1:nS, each = nS), dateYM = rep(1:(nT - 1), each = nS*nS))) %>%
    mutate(year = years[dateYM],
           month = months[dateYM], 
           river = "main")
```

```{r}
#phi
ggplot(phi_tt_main, aes(month, med, color = factor(year))) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = lo, ymax = hi), 
                width = 0.2,
                position = position_dodge(0.05),
                size = 0.75) +
  ylab("Probability of survival +/- 95% CI") +
  xlab("Month") +
  ggtitle("Main - rows = from state, cols = to state") +
  facet_grid(rows = vars(state))

# p
ggplot(p_tt_main, aes(month, med, color = factor(year))) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = lo, ymax = hi), 
                width = 0.2,
                position = position_dodge(0.05),
                size = 0.75) +
  ylab("Probability of detection +/- 95% CI") +
  xlab("Month") +
  ggtitle("Main") +
  facet_grid(rows = vars(state))

#psi
ggplot(psi_tt_main, aes(month, med, color = factor(year))) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = lo, ymax = hi), 
              width = 0.2,
              position = position_dodge(0.03),
              size = 0.75) +
  ylab("Probability of changing state +/- 95% CI") +
  xlab("Month") +
  ggtitle("Main") +
  facet_grid(rows = vars(stateFrom),
             cols = vars(stateTo))
```

## phiT_pT_psiT_trib

Load data 'trib' for analysis

```{r getTargets trib}
#d_tt <- tar_read(target_d_trib)
eh_trib <- tar_read(target_eh_trib)

str(eh_trib$eh)
#kable(eh_trib$eh[1:8,1:10])

table(paste(eh_trib$first, eh_trib$last, sep="_"))

```

Plot trib model estimates

```{r tt trib, eval=TRUE}
### Read the model run into global memory
if (tar_exist_objects(c("tt_modelOut_trib"))) {
  mod_tt_trib <- tar_read(tt_modelOut_trib)
  
  MCMCplot(object = mod_tt_trib$mcmc, params = "betaPhi")
  MCMCplot(object = mod_tt_trib$mcmc, params = "betaP")

   MCMCtrace(object = mod_tt_trib$mcmc,
             ISB = FALSE,
             exact = TRUE,
             params = c("betaPhi[1, 1]", "betaPhi[2, 1]", "betaPhi[3, 1]",
                        "betaPhi[1, 2]", "betaPhi[2, 2]", "betaPhi[3, 2]",
                        "betaPhi[1, 3]", "betaPhi[2, 3]", "betaPhi[3, 3]"),
             pdf = FALSE,
             priors = priors
             )

    MCMCtrace(object = mod_tt_trib$mcmc,
           ISB = FALSE,
           exact = TRUE,
           params = c("betaP[1, 1]", "betaP[2, 1]", "betaP[3, 1]",
                      "betaP[1, 2]", "betaP[2, 2]", "betaP[3, 2]",
                      "betaP[1, 3]", "betaP[2, 3]", "betaP[3, 3]"),
           pdf = FALSE,
           priors = priors
           )

    modSummary_tt_trib <- MCMCsummary(object = mod_tt_trib$mcmc, round = 3) %>%
      rename(lo = '2.5%', med = '50%', hi = '97.5%')
}
```

```{r}
  nS <- mod_tt_trib$myConstants$nStates
  nT <- mod_tt_trib$myConstants$T 
  
  phi_tt_trib <- modSummary_tt_trib %>% 
    filter(substr(row.names(modSummary_tt_trib), 1, 7) == "betaPhi") %>%
    add_column(data.frame(state = 1:nS, dateYM = rep(1:(nT - 1), each = nS))) %>%
    mutate(year = years[dateYM],
           month = months[dateYM], 
           river = "trib")
  
  p_tt_trib <- modSummary_tt_trib %>% 
    filter(substr(row.names(modSummary_tt_trib), 1, 6) == "betaP[") %>%
    add_column(data.frame(state = 1:nS, dateYM = rep(1:(nT - 1), each = nS))) %>%
    mutate(year = years[dateYM],
           month = months[dateYM], 
           river = "trib")

  psi_tt_trib <- modSummary_tt_trib %>% 
    filter(substr(row.names(modSummary_tt_trib), 1, 3) == "psi") %>%
    add_column(data.frame(stateFrom = 1:nS, stateTo = rep(1:nS, each = nS), dateYM = rep(1:(nT - 1), each = nS*nS))) %>%
    mutate(year = years[dateYM],
           month = months[dateYM], 
           river = "trib")
```

```{r}
#phi
ggplot(phi_tt_trib, aes(month, med, color = factor(year))) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = lo, ymax = hi), 
                width = 0.2,
                position = position_dodge(0.05),
                size = 0.75) +
  ylab("Probability of survival +/- 95% CI") +
  xlab("Month") +
  ggtitle("Trib") +
  facet_grid(rows = vars(state))

# p
ggplot(p_tt_trib, aes(month, med, color = factor(year))) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = lo, ymax = hi), 
                width = 0.2,
                position = position_dodge(0.05),
                size = 0.75) +
  ylab("Probability of detection +/- 95% CI") +
  xlab("Month") +
  ggtitle("Trib") +
  facet_grid(rows = vars(state))

#psi
ggplot(psi_tt_trib, aes(month, med, color = factor(year))) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = lo, ymax = hi), 
              width = 0.2,
              position = position_dodge(0.03),
              size = 0.75) +
  ylab("Probability of changing state +/- 95% CI") +
  xlab("Month") +
  ggtitle("Trib - rows = from state, cols = to state") +
  facet_grid(rows = vars(stateFrom),
             cols = vars(stateTo))
```

## Combine main and trib estimates

```{r}
phi_tt_mainTrib <- add_row(phi_tt_main, phi_tt_trib)
psi_tt_mainTrib <- add_row(psi_tt_main, psi_tt_trib)
```

```{r}
ggplot(phi_tt_mainTrib, aes(month, med, color = factor(river))) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = lo, ymax = hi), 
                width = 0.2,
                position = position_dodge(0.05),
                size = 0.75) +
  ylab("Probability of survival +/- 95% CI") +
  xlab("Month") +
  ggtitle("Main and trib") +
  facet_grid(rows = vars(state),
             cols = vars(year))
```

## Calculate means for broad comparisons

```{r}
  # Overall means across occasions
# Main
  (
    phi_tt_main_means <- phi_tt_main %>%
    group_by(state) %>%
    summarise(meanPhi = mean(mean),
              sdPhi = sd(mean),
              meanLo = mean(lo),
              meanMed = mean(med),
              meanHi = mean(hi)) %>%
    mutate(river = "main") %>%
    ungroup()
  )
    
  (
    psi_tt_main_means <- psi_tt_main %>%
    group_by(stateFrom, stateTo) %>%
    summarise(meanPhi = mean(mean),
              sdPhi = sd(mean),
              meanLo = mean(lo),
              meanMed = mean(med),
              meanHi = mean(hi)) %>%
    mutate(river = "main") %>%
    ungroup()
    )

# Trib
  (
    phi_tt_trib_means <- phi_tt_trib %>%
    group_by(state) %>%
    summarise(meanPhi = mean(mean),
              sdPhi = sd(mean),
              meanLo = mean(lo),
              meanMed = mean(med),
              meanHi = mean(hi)) %>%
    mutate(river = "trib") %>%
    ungroup()
    )
    
  (psi_tt_trib_means <- psi_tt_trib %>%
    group_by(stateFrom, stateTo) %>%
    summarise(meanPhi = mean(mean),
              sdPhi = sd(mean),
              meanLo = mean(lo),
              meanMed = mean(med),
              meanHi = mean(hi)) %>%
    mutate(river = "trib") %>%
    ungroup()
    )


```

## Combine mean main and trib estimates

```{r}
phi_tt_mainTrib_means <- add_row(phi_tt_main_means, phi_tt_trib_means)
psi_tt_mainTrib_means <- add_row(psi_tt_main_means, psi_tt_trib_means)
```

Probability of survival for each size state

```{r}
ggplot(phi_tt_mainTrib_means, aes(state, meanMed, color = river)) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = meanLo, ymax = meanHi), 
                width = 0.2,
                position = position_dodge(0.05),
                size = 0.75) +
  ylab("Probability of survival +/- 95% CI") +
  xlab("Size state")
```

Probability of changing state. 'From' size states are on the x-axis, 'to' size states are in the rows of the facets

```{r}
ggplot(psi_tt_mainTrib_means, aes(stateFrom, meanMed, color = river)) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = meanLo, ymax = meanHi), 
              width = 0.2,
              position = position_dodge(0.03),
              size = 0.75) +
  ylab("Probability of changing state +/- 95% CI") +
  xlab("Size state") +
  facet_grid(rows = vars(stateTo))
```

Matrix model

```{r}

# mMain <- matrix(
#   c()
# )
```

## phiT_pT_psiT_mainTrib

Main and trib modeled together. Not using this as of 7-27-22.

```{r ttt, eval=TRUE}
### Read the model run into global memory
# if (tar_exist_objects(c("ttt_modelOut"))) {
#   mod_ttt <- tar_read(ttt_modelOut)
#   
#   #MCMCplot(object = mod$mcmc)
#   
#   modSummary_ttt <- MCMCsummary(object = mod_ttt$mcmc, round = 3)
# }
#kable(modSummary %>%
#        add_column(data.frame(year = rep(years[1:15], 2), dateYM = rep(occs[1:15], 2)) )


#  d %>%
#  summarize(unique(data.frame(dateYM, occ)))

```

```{r ttt phi Summary, eval=TRUE}
# nS <- tar_read(ttt_nStates)
# nT <- tar_read(ttt_myConstants)$T
# 
# modSummaryPhi_ttt <- modSummary_ttt %>%
#   filter(substr(row.names(modSummary), 1, 10) == "betaPhiOut") %>%
#   add_column(data.frame(state = 1:nS, dateYM = rep(1:(nT - 1), each = nS))) %>%
#   mutate(mainTrib = ifelse(state < 4, "Main", "Trib"),
#          size = paste0("Size", (state - 1) %% 3 + 1))
# 
# ggplot(modSummaryPhi_ttt, aes(dateYM, mean)) +
#   geom_point() +
#   geom_line() +
#   facet_grid(mainTrib ~ size)
# 
# 
# # modSummaryYears <- modSummary %>%
# #   filter(substr(row.names(modSummary), 1, 3) == "betaPhiout") %>%
# #   add_column(data.frame(year = years[1:15], dateYM = occs[1:15], occ = 1:15)) %>%
# #   group_by(year) %>%
# #   mutate(maxSampPerYear = occ == max(occ))
# # 
# # kable(
# #   modSummaryYears %>%
# #   group_by(year) %>%
# #   filter(!maxSampPerYear) %>%
# #   summarize(phiProd = prod(mean),
# #             dateRange = range(dateYM)) %>%
# #     as.data.frame()
# # )
#   MCMCplot(object = mod$mcmc, params = "betaPhiRiverOut")
# 
#   priors <- rnorm(tar_read(ttt_runData)$nIter * tar_read(ttt_runData)$nChains, 0, 1/sqrt(.1))
#   MCMCtrace(object = mod$mcmc,
#             #ISB = FALSE,
#             #exact = TRUE, 
#             params = c("betaPhiRiverOut"),
#             pdf = FALSE, 
#             priors = priors
#             )
#   
#     MCMCtrace(object = mod$mcmc,
#             #ISB = FALSE,
#             #exact = TRUE, 
#             params = c("betaPhiOut"),
#             pdf = FALSE, 
#             priors = priors
#             )

```

```{r p summary ttt, eval=TRUE}
#create data frame for summarizing p results

# modSummaryYearsP <- modSummary %>%
#   filter(substr(row.names(modSummary), 1, 2) == "p[") %>%
#   add_column(data.frame(year = years[1:15], dateYM = occs[1:15], occ = 1:15)) %>%
#   group_by(year) %>%
#   mutate(maxSampPerYear = occ == max(occ)) 
# 
# kable(
#   modSummaryYearsP %>%
#   group_by(year) %>%
#   summarize(pMean = mean(mean),
#             dateRange = range(dateYM))
#   )
```

```{r psi summary ttt}

# # modSummaryYearsP <- modSummary %>%
# #   filter(substr(row.names(modSummary), 1, 2) == "p[") %>%
# #   add_column(data.frame(year = years[1:15], dateYM = occs[1:15], occ = 1:15)) %>%
# #   group_by(year) %>%
# #   mutate(maxSampPerYear = occ == max(occ))
# 
# modSummaryPsi_ttt <- modSummary_ttt %>%
#   filter(substr(row.names(modSummary_ttt), 1, 3) == "psi") %>%
#   add_column(data.frame(state = 1:nS, state2 = rep(1:nS, each = nS), dateYM = rep(1:(nT - 1), each = nS * nS))) %>%
#   mutate(mainTrib = ifelse(state < 4, "Main", "Trib"),
#          size = paste0("Size", (state - 1) %% 3 + 1))
# 
# ggplot(modSummaryPsi_ttt, aes(dateYM, mean, color = factor(state2))) +
#   geom_point() +
#   facet_grid(mainTrib ~ size)
# 

```
