---
title: "Delaware River PIT tag data analysis"
author: "Ben Letcher"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

library(tidyverse)
library(lubridate)
library(knitr)
library(targets)
library(tarchetypes)
library(MCMCvis)
library(htmltools)

library(bs4Dash)
library(DT)
library(gt)
library(pingr)
library(shinybusy)
library(shinyWidgets)

options(tinytex.verbose = TRUE)
```
*Note*: the results shown here are preliminary and have not been officially reviewed by USGS, NYDEC or PA Fish and Boat  

This notebook uses [targets](https://books.ropensci.org/targets/) to manage running code and updating R objects. Targets sets up dependencies among specified objects and only re-runs code as necessary (when an upstream component gets updated). This can save run times for projects with models that take a while to run, like capture-mark-recapture models.  

Data preparation and model running happens using targets and exploration of the data and model runs is below in this Markdown document.  

'knit' the document to update all targets and the markdown exploration below.  

'tar_make()' runs all the R scripts and functions specified in '_targets.R'. Only updated code or sections that are downstream from updated data are re-run.  
'tar_read()' reads 'target' data into the global environment.  

This section (tar_make()) reruns the model   
```{r tar_make}
#  tar_watch(seconds = 10, outdated = FALSE, targets_only = TRUE)


# comment this out when kniting - get Latex error that it can't find the check mark the tar_make() uses
#tar_make()

# tar_prune() # cleans unused data files
#tar_invalidate(everything())
#tar_invalidate(ends_with("ttt"))


#str(d)
```

Load data for analysis
```{r getTargets}
dRaw0 <- tar_read(dRaw0) #all data - including untagged
dRaw <- tar_read(dRaw) #all data for CMR models
d <- tar_read(target_d)
eh <- tar_read(target_eh)

```


Visualize the network - does not work with pdf output
```{r}
#tar_visnetwork()
```

Which rivers (Water) riverN corresponds to
```{r riverN}
table(d$Water, d$riverN)
```

### Raw data summary tables
```{r basic summaries}

kable(data.frame(ftable(d$date)))
#kable(data.frame(ftable(d$Water, d$riverN)))

#kable(data.frame(ftable(d$Water, d$riverN, d$date)))
kable(data.frame(ftable(d$species)))

### Number of unique tags
length(unique(d$tag))
```

Group observations by month.  
Luckily, sampling periods do not span months, so we can use month as a grouping variable for sampling occasion
```{r by month}
kable(data.frame(ftable(d$dateYM)))

```

### Tag information  
Grouped by Water (sampling area)
```{r tag information}

tagN <- d %>% 
  group_by(tag, Water) %>%
  summarize(n = n()) %>%
  filter(tag != "") %>%
  arrange(desc(n))

### Number of times individual fish were observed
table(tagN$n)

### Number of times individual fish were observed by river
(table(tagN$Water, tagN$n))



```

Grouped by main/trib
```{r tag information mainTrib}

tagN_mt <- d %>% 
  group_by(tag, mainTrib) %>%
  summarize(n = n()) %>%
  filter(tag != "") %>%
  arrange(desc(n))

### Number of times individual fish were observed
table(tagN_mt$n)

### Number of times individual fish were observed by river
(table(tagN_mt$mainTrib, tagN_mt$n))


```

### Basic summary plots of raw tagging data
```{r basic graphs}

ggplot(d %>% filter(!is.na(species)), aes(Length)) +
  geom_freqpoly() +
  facet_grid(species ~ Water)

```

```{r data for TAME, ignore=TRUE}

# dTame <- d %>% 
#             select(Latitude, Longitude, tag, dateTime, species, Length, Weight) %>%
#             filter(tag != "", tag != "ad")
#             
# write.csv(dTame, './dataOut/dTame.csv', row.names = FALSE)

```



### Encounter histories  
This is the data structure for the capture-recapture models. Each column is a sampling 'occasion' (here = month) and each row is an individual, where a '1' indicates capture and a '0' indicates not captured.
```{r eh}

str(eh$eh)
kable(head(eh$eh,8))

table(paste(eh$first, eh$last, sep="_"))
```

Summary info for years and occasions
```{r}

years <- colnames(eh$eh) %>%
  substr(6,9) %>%
  as.numeric()

occs <- colnames(eh$eh)
```


### Models
'phi' = apparent survival (probability of staying in the area = p(survival) + p(not moving out of area)).  
'p' = probability of capture given that the fish is alive.
```{r ttt, eval=TRUE}
### Read the model run into global memory
mod <- tar_read(ttt_modelOut)

#MCMCplot(object = mod$mcmc)

modSummary <- MCMCsummary(object = mod$mcmc, round = 3)

#kable(modSummary %>%
#        add_column(data.frame(year = rep(years[1:15], 2), dateYM = rep(occs[1:15], 2)) )


#  d %>%
#  summarize(unique(data.frame(dateYM, occ)))

```

```{r phi Summary, eval=TRUE}
nS <- tar_read(nStates)
nT <- tar_read(ttt_myConstants)$T

modSummaryPhi <- modSummary %>%
  filter(substr(row.names(modSummary), 1, 10) == "betaPhiOut") %>%
  add_column(data.frame(state = 1:nS, dateYM = rep(1:(nT - 1), each = nS))) %>%
  mutate(mainTrib = ifelse(state < 4, "Main", "Trib"),
         size = paste0("Size", (state - 1) %% 3 + 1))

ggplot(modSummaryPhi, aes(dateYM, mean)) +
  geom_point() +
  geom_line() +
  facet_grid(mainTrib ~ size)


# modSummaryYears <- modSummary %>%
#   filter(substr(row.names(modSummary), 1, 3) == "betaPhiout") %>%
#   add_column(data.frame(year = years[1:15], dateYM = occs[1:15], occ = 1:15)) %>%
#   group_by(year) %>%
#   mutate(maxSampPerYear = occ == max(occ))
# 
# kable(
#   modSummaryYears %>%
#   group_by(year) %>%
#   filter(!maxSampPerYear) %>%
#   summarize(phiProd = prod(mean),
#             dateRange = range(dateYM)) %>%
#     as.data.frame()
# )
  MCMCplot(object = mod$mcmc, params = "betaPhiRiverOut")

  priors <- rnorm(tar_read(ttt_runData)$nIter * tar_read(ttt_runData)$nChains, 0, 1/sqrt(.1))
  MCMCtrace(object = mod$mcmc,
            #ISB = FALSE,
            #exact = TRUE, 
            params = c("betaPhiRiverOut"),
            pdf = FALSE, 
            priors = priors
            )

```


```{r p summary, eval=TRUE}
#create data frame for summarizing p results

# modSummaryYearsP <- modSummary %>%
#   filter(substr(row.names(modSummary), 1, 2) == "p[") %>%
#   add_column(data.frame(year = years[1:15], dateYM = occs[1:15], occ = 1:15)) %>%
#   group_by(year) %>%
#   mutate(maxSampPerYear = occ == max(occ)) 
# 
# kable(
#   modSummaryYearsP %>%
#   group_by(year) %>%
#   summarize(pMean = mean(mean),
#             dateRange = range(dateYM))
#   )
```

```{r psi summary}

# modSummaryYearsP <- modSummary %>%
#   filter(substr(row.names(modSummary), 1, 2) == "p[") %>%
#   add_column(data.frame(year = years[1:15], dateYM = occs[1:15], occ = 1:15)) %>%
#   group_by(year) %>%
#   mutate(maxSampPerYear = occ == max(occ))

modSummaryPsi <- modSummary %>%
  filter(substr(row.names(modSummary), 1, 3) == "psi") %>%
  add_column(data.frame(state = 1:nS, state2 = rep(1:nS, each = nS), dateYM = rep(1:(nT - 1), each = nS * nS))) %>%
  mutate(mainTrib = ifelse(state < 4, "Main", "Trib"),
         size = paste0("Size", (state - 1) %% 3 + 1))

ggplot(modSummaryPsi, aes(dateYM, mean, color = factor(state2))) +
  geom_point() +
  facet_grid(mainTrib ~ size)


```

